CES analysis for Paul Brockway rebound paper

### 2015-07-30 Articles/Brockway_Rebound_2015/Brockway_2015.Rnw
New files for Brockway analysis.

\journal{Ecological Economics}

@@

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
library(EconData)
library(EconModels)
library(dplyr)
library(reshape2)
library(ggplot2)
library(mosaic)
library(xtable)
library(knitr)
library(texreg)

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
# ********************************************************
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)

# This function is used in many places below.
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE, 
        booktabs=TRUE)
}
@

@@

\title{Energy and the CES production function}

@@

\author[IST]{Jo\~{a}o dos Santos}
\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\author[Leeds]{Paul Brockway}
\ead{mkh2@calvin.edu}
\author[Leeds]{Marco Sekai}
\author[IST]{Tiago Domingos}
\author[CalvinMath]{Randall Pruim}

\address[IST]{IST, Lisbon, Portugal}
\address[Leeds]{Leeds University, Leeds, UK}
\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\address[CalvinMath]{Mathematics \& Statistics Department, Calvin College, Grand Rapids, MI 49546, USA}
\cortext[cor1]{Corresponding author}

\begin{abstract}
%% Text of abstract
**** Add abstract ****
\end{abstract}

@@

This is a study of the effects of quality-adjusted production function inputs
on the importance of energy in the production function.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Coordinates of Analysis} 
\label{sec:Coordinates}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes the coordinates of analysis and briefly reviews 
literature related to each.

%++++++++++++++++++++++++++++++
\subsection{Mathematical Forms of the Energy-augmented Production Function} 
\label{sec:Prod_Func_Forms}
%++++++++++++++++++++++++++++++

In this paper, we assess a prominent energy-augmented production function:
Constant Elasticity of Substitution (CES).
The CES model is assessed relative to 
a model of exponential growth only.
The following subsections describe each.

%------------------------------
\subsubsection{Constant Elasticity of Substitution Production Function (CES)} 
\label{sec:CES}
%------------------------------

The energy-augmented 
Constant Elasticity of Substitution (CES) production function%
  \footnote{
  We use the The \texttt{R} 
  % \citep{R} 
  package \texttt{micEconCES} 
  % \citep{Henningsen:2011td} 
  to estimate the CES production function.
  }
can be given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$y \equiv Y/Y_{0}$,
$\gamma$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is the Solow residual (often thougth to represent the pace of technological progress),
$t$ (time) is measured in years,
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$),
$k \equiv K/K_{0}$, 
$l \equiv L/L_{0}$, 
$Y$ (economic output) is represented by GDP, 
$K$ (capital) is expressed in currency units, 
$L$ (labor) is expressed in workers or work-hours/year, and
the 0 subscript indicates values at an initial year.%
  \footnote{Dimensionless, indexed quantities are represented by 
  lower-case symbols 
  ($y$, $k$, $l$, $e$, $q$, $x$, and $u$), and dimensional 
  quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, $E$, $Q$, $X$, and $U$). 
  Model parameters are represented by Greek letters
  ($\alpha_1$, $\alpha_2$, $\lambda$, $\theta$).
  }
The CES model without energy is given in Equation~\ref{eq:CESkl}.
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)} . 
\end{equation}
%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy 
using a ($kl$)($e$) nesting structure, as is typical in the literature. 
Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$. 
%
\begin{equation} \label{eq:CESkle}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}.
\end{equation}
%
In the CES production function, 
$\gamma$ is a fitting parameter that accounts for an atypical first year. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).
The elasticity of substitution 
between capital ($k$) and labor ($l$) is given by 
$\sigma_1 = \frac{1}{1+\rho_1}$, and
the elasticity of substitution between ($kl$) and ($e$) is given by 
$\sigma = \frac{1}{1+\rho}$. 
As $\rho_1 \rightarrow 0$, $\sigma_1 \rightarrow 1$,
and the embedded CES production function for $k$ and $l$ degenerates 
to the Cobb-Douglas production function.
Similarly, as $\rho \rightarrow 0$, $\sigma \rightarrow 1$,
and the CES production function for $(kl)$ and $(e)$ degenerates 
to the Cobb-Douglas production function.
As $\sigma \rightarrow \infty$ ($\rho \rightarrow -1$), 
($kl$) and ($e$) are perfect substitutes. 
As $\sigma \rightarrow 0$ ($\rho \rightarrow \infty$), 
($kl$) and ($e$) are perfect complements: 
no substitution is possible. Similarly, 
as $\sigma_1 \rightarrow 0$ ($\rho_1 \rightarrow \infty$),
$k$ and $l$ are perfect complements.
$\delta_1$ describes the relative importance of capital ($k$)
and labor ($l$), and
$\delta$ describes the importance of ($kl$) relative to ($e$).

Constraints on the fitting parameters include 
$\delta_1 \in [0,1]$,
$\delta \in [0,1]$,
$\rho_1 \in [-1,0) \cup (0,\infty)$, and
$\rho \in [-1,0) \cup (0,\infty)$.

Two other nestings of the factors of production ($k$, $l$, and $e$)
are possible with the CES model.
%
\begin{equation} \label{eq:CESlek}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1} 
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
%
\begin{equation} \label{eq:CESekl}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1} 
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}
\end{equation}
% 
Note that $\rho$ ($\sigma$), $\rho_1$ ($\sigma_1$), $\delta$, and $\delta_1$ have different
meanings depending upon the nesting of the factors of production.

For the CES model, factors shares~($\alpha_1$, $\alpha_2$, and $\alpha_3$) are calculated by
$\alpha_1 = \delta_1$, $\alpha_2 = 1 - \delta_1$, and $\alpha_3 = 0$ 
for the CES model with two factors of production
and by 
$\alpha_1 = \delta \delta_1$, $\alpha_2 = \delta(1 - \delta_1)$, and $\alpha_3 = 1 - \delta$
for the CES model with three factors of production and the ($kl$)$e$ nesting.
The values of $\alpha_1$, $\alpha_2$, and $\alpha_3$ are interpreted as
factor shares for capital, labor, and energy, respectively,
as shown in Table \ref{tab:CES_abg_equations}.

\begin{table} \caption{Equations for $\alpha_1$, $\alpha_2$, and $\alpha_3$ 
                        (factor shares for capital, labor, and energy, respectively)
                        for the various CES nestings.} 
\label{tab:CES_abg_equations} 
  \begin{center}
    \begin{tabular}{c c c c} 
      \toprule
      Nesting    & $\alpha_1$              & $\alpha_2$              & $\alpha_3$       \\
      \midrule
      ($kl$) \;  & $\delta_1$              & $1 - \delta_1$          & 0          \\
      ($kl$)$e$  & $\delta \delta_1$       & $\delta (1 - \delta_1)$ & $1 - \delta$ \\   
      ($le$)$k$  & $1 - \delta$            & $\delta \delta_1$       & $\delta (1 - \delta_1)$ \\
      ($ek$)$l$  & $\delta (1 - \delta_1)$ & $1 - \delta$            & $\delta \delta_1$ \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

%------------------------------
\subsubsection{Exponential Production Function (reference model)} 
\label{sec:exp}
%------------------------------

We define an exponential-only reference model for economic growth
in Equation~\ref{eq:exponential-only-model}.
%
\begin{equation} \label{eq:exponential-only-model}
  y = \theta A \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)} \; .
\end{equation}
%
We expect that the reference model
will have a larger fitted Solow residual term than
the CES model, because no factors of production are
included in the reference model to drive economic growth.
Indeed, in the reference model, 
all economic growth is attributed to the Solow residual~($\lambda$).

In contrast, it is not necessarily true that 
CES models will exhibit lower $mse$ than the reference model
shown in Equation~\ref{eq:exponential-only-model}.
The CES model has more fitting parameters,
but it incorporates the factors of production
at constant returns to scale. 
The reference model has $\alpha_1 + \alpha_2 + \alpha_3 = 0$,
whereas the CES model~(Equation~\ref{eq:CESgeneric})
implicitly assumes constant returns to scale.
Thus, if the factors of production~($\alpha_1$, $\alpha_2$, and $\alpha_3$) 
are poorly correlated to output~($y$),
the constant returns to scale constraint may cause higher $mse$ for
the CES model relative to the reference model.

%++++++++++++++++++++++++++++++
\subsection{Economies} 
\label{sec:Economies}
%++++++++++++++++++++++++++++++

Discuss economies here.

UK and Portugal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sources of Data}
\label{sec:Sources_of_Data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Discuss data sources here.

%++++++++++++++++++++++++++++++
\subsection{Historical Data} 
\label{sec:Historical_Data}
%++++++++++++++++++++++++++++++

Historical data are stored in the \texttt{IST} and \texttt{Leeds} data sets.

<<Formulas>>=
# Set up the formulas for each type of analysis.
# These formulas contain energy. 
# We can use keep_summands to drop the energy term (3rd summand on the rhs)
# later if needed.
formulas <- 
  list( 
    PT = c(unadjusted = build_formula("iGDP", list("iKstkS.L", "iL", "iXpMP", "iYear")),
             adjusted = build_formula("iGDP", list("iKservS.L", "ihLest", "iUMP", "iYear"))),
    UK = c(unadjusted = build_formula("iGDP", list("iKstkO.WwithRD", "iL", "iXp", "iYear")),
             adjusted = build_formula("iGDP", list("iKservO.WwithRD", "ihLest", "iU", "iYear")))
  )
countries <- names(formulas)
flavors <- names(formulas$PT)
@

<<Data>>=
# Create the historicalData data frames
historicalData <- list()
commonVars <- c("Year", "Country", "Source")
historicalData$PT <- filter(IST, Country == "PT") 
historicalData$PT <- historicalData$PT[ , unique(c(commonVars,
                                                   all.vars(formulas$PT$unadjusted), 
                                                   all.vars(formulas$PT$adjusted)))]
historicalData$UK <- filter(Leeds, Country == "UK")
historicalData$UK <- historicalData$UK[ , unique(c(commonVars,
                                                   all.vars(formulas$UK$unadjusted), 
                                                   all.vars(formulas$UK$adjusted)))]
@

For unadjusted variables, we use the time series shown in Table~\ref{tab:unadjusted_variables}.

\begin{table} \caption{Unadjusted time-series variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$unadjusted, n=1, right=FALSE)}   
                                                             & \Sexpr{summands(formulas$UK$unadjusted, n=1, right=FALSE)}\\
      Capital     & \Sexpr{summands(formulas$PT$unadjusted, 1)} & \Sexpr{summands(formulas$UK$unadjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$unadjusted, 2)} & \Sexpr{summands(formulas$UK$unadjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$unadjusted, 3)} & \Sexpr{summands(formulas$UK$unadjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

For quality-adjusted variables, we use the time series shown in Table~\ref{tab:quality_adjusted_variables}.

\begin{table} \caption{Quality-adjusted time series variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{summands(formulas$PT$adjusted, n=1, right=FALSE)}   
                                                           & \Sexpr{summands(formulas$UK$adjusted, n=1, right=FALSE)} \\
      Capital     & \Sexpr{summands(formulas$PT$adjusted, 1)} & \Sexpr{summands(formulas$UK$adjusted, 1)} \\
      Labor       & \Sexpr{summands(formulas$PT$adjusted, 2)} & \Sexpr{summands(formulas$UK$adjusted, 2)} \\
      Energy      & \Sexpr{summands(formulas$PT$adjusted, 3)} & \Sexpr{summands(formulas$UK$adjusted, 3)} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

Unadjusted and quality-adjusted data for both Portugal and the United Kingdom 
are shown in Figure~\ref{fig:Historical Data Graph}.

<<Historical Data Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
# Melt the data into a form that is useful in a couple places
histDataForGraph <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  measure.vars <- c(summands(form, right=FALSE), summands(form, 1:3))
  trimmed_historical_data <- historicalData[[Country]][c("Year", measure.vars)]
  mtrimmed_historical_data <- melt(trimmed_historical_data, measure.vars = measure.vars, 
                                   value.name = "value", variable.name = "variable")
  mtrimmed_historical_data$Country <- Country
  mtrimmed_historical_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mtrimmed_historical_data)
}) %>% do.call(rbind, .)

# Relevel factors before plotting
histDataForGraph$Country <- relevelFactor(as.factor(histDataForGraph$Country), c("PT", "UK"))
histDataForGraph$flavor <- relevelFactor(as.factor(histDataForGraph$flavor), c("Unadjusted", "Quality-adjusted"))
histDataForGraph$variable <- relevelFactor(as.factor(histDataForGraph$variable), unique(histDataForGraph$variable))

historicalPlot(histDataForGraph, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(1,2,4,3,2,4,3,2,3,2,3)) + 
  scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
  coord_cartesian(ylim=c(0.5, 13)) +
  xlab("") + 
  ylab("Indexed (1960=1)")
@

<<Pre-econometric graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Pre-econometric data.", warning=FALSE>>=
pre_econometric_data <- leaf_apply(formulas, class = "formula", f=function(form, name){
  meta <- strsplit(name, split="[.]") %>% unlist
  Country <- meta[1]
  flavor <- meta[2]
  nums <- summands(form, n=1:3)
  denom <- summands(form, n=1, right=FALSE)
  ratios <- lapply(nums, function(num) {paste(num, denom, sep="/")}) %>% unlist
  coltemplate <- build_formula(NULL, c("Year", ratios))
  trimmed_historical_data <- historicalData[[Country]][all.vars(coltemplate)]
  ratio_data <- evalFormula(coltemplate, data=trimmed_historical_data)$right
  mratio_data <- melt(ratio_data, measure.vars = ratios, value.name = "ratio", variable.name = "variable")
  mratio_data$Country <- Country
  mratio_data$flavor <- if (flavor == "adjusted") "Quality-adjusted" else "Unadjusted"
  return(mratio_data)
}) %>% do.call(rbind, .)

# Fix levels
pre_econometric_data$flavor <- relevelFactor(as.factor(as.character(pre_econometric_data$flavor)), 
                                               c("Unadjusted", "Quality-adjusted"))
pre_econometric_data$Country <- relevelFactor(as.factor(pre_econometric_data$Country), 
                                              c("PT", "UK"))
pre_econometric_data$variable <- relevelFactor(as.factor(pre_econometric_data$variable),
                                               unique(pre_econometric_data$variable))
# Now make the plot of pre-econometric data
historicalPlot(pre_econometric_data, 
               mapping=aes(x=Year, y=ratio, group=variable, linetype=variable), 
               facet_formula = Country ~ flavor, 
               line_types = c(2,4,3, 2,4,3, 2,3,2, 3)) + 
  xlab("") + 
  ylab("Indexed (1960=1)")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation (Without Cost-share Theorem)}
\label{sec:parameter_estimation_noCST}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Parameter estimation is accomplished with \texttt{formula}s of the form
% %
% \begin{equation} \label{eq:formulas}
%   y \sim x_1 + x_2 + x_3 + time,
% \end{equation}
% %
% where $y$ is economic output and 
% $x_1$, $x_2$, and $x_3$ are factors of production
% and permutations of
% capital ($k$), labor ($l$), and energy ($e$).
% (The factors of production may be quality-adjusted.)
% $time$ is the time variable, usually \texttt{iYear}. 
% The \texttt{formula}s are used in a fitting function such as 
% \texttt{cdModel} to create a \texttt{model} object.
% The function \texttt{naturalCoef} extracts reasonable
% coefficients from \texttt{model} objects.
% 
% Note that in the fitting process for the Cobb-Douglas equation,
% $\texttt{scale} = \theta$, 
% $\alpha_1 = \alpha$,
% $\alpha_2 = \beta$, and
% $\alpha_3 = \gamma$,
% when the \texttt{formula} is specified as 
% %
% \begin{equation} \label{eq:formula_kle}
%   \texttt{y} \sim \texttt{k + l + e + time}.
% \end{equation}
% %
% Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity.

The models being fit can be described by the algebraic form of the model 
(Cobb-Douglas, CES, etc.) and a formula that enumerates 
which data variables are being used in which roles to fit the model.  
A formula of the form
%
\begin{equation} \label{eq:formulas-3-factor}
y \sim x_1 + x_2 + x_3 + t,
\end{equation}
%
or 
%
\begin{equation} \label{eq:formulas-2-factor}
y \sim x_1 + x_2 + t,
\end{equation}
%
describes the economic output variable ($y$, usually \texttt{iGDP},
                                        indexed GDP) 
and the factors of production ($x_1$, $x_2$, and $x_3$,
                               which will be some measure of capital, labor, and energy, 
                               but perhaps not in that order), 
and a time variable ($t$, usually \texttt{iYear}, 
                     the number of years since the beginning of data collection).  
All of the models assume an error term that is additive on the 
logarithmic scale and are fit by the method of least squares.  
Model fitting provides estimates for all parameters in the model.

<<Models>>=
fileNameCDCES <- "EcolEcon2015ModelsCDCES.rds"
fileNameCST <- "EcolEcon2015ModelsCST.rds"
fileNameExp <- "EcolEcon2015ModelsExp.rds"
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(fileNameCDCES)
  modelsCST <- readRDS(fileNameCST)
  modelsExp <- readRDS(fileNameExp)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.

  #
  # Models that are exponential only
  #
  modelsExp <- list()
  alphaExp <- 0
  betaExp <- 0
  gammaExp <- 0
  for (country in countries){
    # Replace with better code later. 
    # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0, alpha_2 = 0, alpha_3 = 0))
    # or something like that.
    form <- keep_summands(formulas[[country]][["unadjusted"]], n = c(-1, -2, -3)) # Remove energy
    yVar <- as.name(summands(form, n=1, right=FALSE))
    tVar <- as.name(summands(form, n=1))
    data <- historicalData[[country]]
    expMod <- lm(log(eval(yVar)) ~ eval(tVar), data=data)
    attr(expMod, "naturalCoeffs") <- data.frame(lambda=coef(expMod)[[2]],
                                                  logscale=coef(expMod)[[1]], 
                                                  scale=exp(coef(expMod)[[1]]),
                                                  alpha_1=alphaExp,
                                                  alpha_2=betaExp, 
                                                  alpha_3=gammaExp)
    expMod$formula <- form
    sdata <- subset(data, select=all.vars(form))
    sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
    expMod$data <- sdata
    expMod$response <- sdata$iGDP 
    modelsExp[[country]] <- expMod
  }
  # Save exponential theorem models to disk.
  saveRDS(modelsExp, file=fileNameExp)

  #
  # Models that adhere to the cost-share theorem
  #

  modelsCST <- list()
  # Output elasticities based on cost share theorem
  alphaCST <- 0.3 # Fraction of national accounts spent on capital
  betaCST <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    for (adj in c("unadjusted", "adjusted")){
      # Replace with better code later. 
      # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0))
      # or something like that.
      form <- keep_summands(formulas[[country]][[adj]], n=-3) # Remove energy
      yVar <- as.name(summands(form, n=1, right=FALSE))
      kVar <- as.name(summands(form, n=1))
      lVar <- as.name(summands(form, n=2))
      tVar <- as.name(summands(form, n=3))
      data <- historicalData[[country]]
      # Cobb-Douglas
      cdModCST <- lm(log(eval(yVar)) - alphaCST*log(eval(kVar)) - betaCST*log(eval(lVar)) ~ eval(tVar), data = data)
      attr(cdModCST, "naturalCoeffs") <- data.frame(lambda=coef(cdModCST)[[2]],
                                                    logscale=coef(cdModCST)[[1]], 
                                                    scale=exp(coef(cdModCST)[[1]]),
                                                    alpha_1=alphaCST,
                                                    alpha_2=betaCST, 
                                                    alpha_3=0)
      cdModCST$formula <- form
      sdata <- subset(data, select=all.vars(form))
      sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
      cdModCST$data <- sdata
      cdModCST$response <- sdata$iGDP
      modelsCST[[country]][[adj]][["noE"]][["CD"]] <- cdModCST
      # CES
      cesModCST <- plm(log(eval(yVar) / (alphaCST*eval(kVar)^(-rho_1) + betaCST*eval(lVar)^(-rho_1))^(-1/rho_1)) ~ eval(tVar),
                       data = historicalData[[country]],
                       params=c(rho_1=0.8))
      names(cesModCST$coefficients) <- c("logscale", "lambda", "rho_1")
      cesModCST$coefficients <- c(coef(cesModCST), c(delta_1=alphaCST, delta=1))
      attr(cesModCST, "naturalCoeffs") <- EconModels:::makeNatCoef(cesModCST)
      cesModCST$formula <- form
      cesModCST$data <- sdata
      modelsCST[[country]][[adj]][["noE"]][["CES"]][["kl"]] <- cesModCST
    }
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCST, file=fileNameCST)

  #
  # Models where all parameters are estimated (models don't adhere to the cost share theorem)
  #

  models <- list()
  for(country in countries){
    for (flavor in flavors){
      fnoE <- keep_summands(formulas[[country]][[flavor]], -3) # Drop the energy term (3rd term on RHS)
      fwithE <- formulas[[country]][[flavor]]
      # CD models
      models[[country]][[flavor]][["noE"]][["CD"]]   <- cdModel(fnoE, data = historicalData[[country]])
      models[[country]][[flavor]][["withE"]][["CD"]] <- cdModel(fwithE, data = historicalData[[country]])
      # CES models
      models[[country]][[flavor]][["noE"]][["CES"]][["kl"]] <- cesModel(fnoE, 
                                                                        data = historicalData[[country]], 
                                                                        nest=c(1,2))
      nestsWithE <- list(kle=c(1,2,3), lek=c(2,3,1), ekl=c(3,1,2))
      models[[country]][[flavor]][["withE"]][["CES"]] <- lapply(nestsWithE, function(nest){
          cesModel(formula=fwithE, data=historicalData[[country]], nest=nest)
        })
    }      
  }
  # Save models to disk.
  saveRDS(models, file=fileNameCDCES)
}
# Put all exponential model coefficients into a data frame
AllCoeffsExp <- leaf_apply(modelsExp, class = c("lm"), f = function(modExp, id){
  # Parse the id to find details of this model
  id_list <- strsplit(id, split = "[.]")[[1]]
  res <- rbind(list(Country = id_list[[1]], 
                    flavor = NA,
                    energy = "Without energy", 
                    model = "exp", 
                    nest = NA,
                    cst = "Reject CST"))
  res <- cbind(res, naturalCoef(modExp))
  # Now caluculate mean squared error (mse) for each model.
  sqResids <- resid(modExp)^2
  mse <- sum(sqResids) / (length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)

# Put all the cost-share theorem (CST) model coefficients into a data frame
AllCoeffsCST <- leaf_apply(modelsCST, class = c("lm"), f = function(modCST, id){
  # Parse the id to find out whether this is quality-adjusted data, etc.
  id_list <- strsplit(id, split = "[.]")[[1]]
  res <- rbind(list(Country = id_list[[1]], 
                    flavor = if (id_list[[2]]=="unadjusted") "Unadjusted" else "Quality-adjusted",
                    energy = if (id_list[[3]]=="noE") "Without energy" else "With energy",
                    model=id_list[[4]], 
                    nest=if(length(id_list) > 4) id_list[[5]] else NA,
                    cst="Adhere to CST"))
  res <- cbind(res, naturalCoef(modCST))
  # Now calculate the mean squared error (mse) for each model.
  sqResids <- resid(modCST)^2
  mse <- sum(sqResids)/(length(sqResids))
  res <- cbind(res, list(mse=mse))
  return(res)
}) %>% do.call(plyr::rbind.fill, .)

>>>>

\journal{Unknown}

@@

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
library(EconData)
library(EconModels)
library(dplyr)
library(reshape2)
library(ggplot2)
# library(mosaic)
# library(xtable)
library(knitr)
# library(texreg)

# Tell whether to re-run the analysis to create the models.
# This file will use previously-saved models to generate
# the paper if set to FALSE, greatly reducing time.
# If something changes in the data or in the code that
# performs the analysis, set 
# rerunModels = TRUE
# once to re-generate the information needed to 
# create the paper.
# ********************************************************
rerunModels=FALSE
# ********************************************************
opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)
@

@@

\title{CES analysis for Paul Brockway rebound paper}

@@

\author[CalvinEngr]{Matthew Kuperus Heun\corref{cor1}}
\ead{mkh2@calvin.edu}

\address[CalvinEngr]{Engineering Department, Calvin College, Grand Rapids, MI 49546, USA}
\cortext[cor1]{Corresponding author}

\begin{abstract}
%% Text of abstract
This document provides an analysis of the CES production function
for UK, US, and CN using quality-adjusted factors of production
and the ($kl$)$e$ nesting.
\end{abstract}

@@

This is a study of the effects of the CES production function 
using quality-adjusted factors of production
and the ($kl$)$e$ nesting.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{CES production function} 
\label{sec:ces}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The version of the CES production function being fitted is:
%
\begin{equation} \label{eq:CESkle}
  y = \theta \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho} \; ; \;
      A \equiv \mathrm{e}^{\lambda (t-t_0)} \; .
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Historical data} 
\label{sec:historical_data}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Historical data have been provided by Paul Brockway.

<<Data>>=
# Create the historicalData data frames
# Set the working directory to the project directory.
print(getwd())
historicalData <- read.csv(file = "data/BrockwayData.csv")
@

<<Historical Data Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Historical data.", warning=FALSE>>=
# Melt data for plotting
historicalDataForGraph <- melt(data = historicalData, measure.vars = c("iY", "iK", "iL", "iU"), 
                               value.name = "value", variable.name = "variable")

# Relevel factors before plotting
historicalDataForGraph$Country <- relevelFactor(as.factor(historicalDataForGraph$Country), c("UK", "US", "CN"))

historicalPlot(historicalDataForGraph, 
               mapping=aes(x=Year, y=value, group=variable, linetype=variable), 
               facet_formula = ~ Country, 
               line_types = c(1,2,4,3)) + 
  scale_y_continuous(breaks=c(1, 10, 20)) +
  xlab("") + 
  ylab("Indexed Quantities")
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Parameter Estimation}
\label{sec:parameter_estimation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

<<Models>>=
fileName <- file.path("models", "Brockway2015Models.rds")
if (!rerunModels){
  # load from the file on disk
  models <- readRDS(fileName)
} else {
  # re-run the models and save to disk.
  # In RStudio, be sure to set the working directory to the source file directory.
  models <- lapply(c("UK", "US", "CN"), function(coun){
    return(cesModel(iY ~ iK + iL + iU + iYear, filter(historicalData, Country==coun, nest=1:3)))
  })
  # Save the models.
  saveRDS(models, file = fileName)
}
@
